Appendix 3: Estimating species interactions

### 2016-02-18 Harris_2015/manuscript-materials/draft.Rmd
many changes

**Abstract:** Inferring species interactions from co-occurrence data is one of
the most controversial tasks in community ecology. One difficulty is that a
single pairwise interaction can ripple through an ecological network and produce
surprising indirect consequences. For example, the negative correlation between
two competing species can be reversed in the presence of a third species that is
capable of outcompeting both of them. Here, I apply models from statistical
physics, called Markov networks or Markov random fields, that can predict
the direct and indirect consequences of any possible species interaction matrix.
Interactions in these models can be estimated from observed co-occurrence
rates via maximum likelihood, controlling for indirect effects. Using simulated
landscapes with known pairwise interaction strengths, I evaluated Markov
networks and six existing approaches. The Markov networks consistently
outperformed other methods, correctly isolating direct interactions between
species pairs even when indirect interactions or abiotic factors largely
overpowered them. Two computationally efficient approximations, based on
controlling for indirect effects with linear or generalized linear models, also
performed well. Indirect effects reliably caused a common null modeling approach
to produce incorrect inferences, however.

@@

\noindent \textbf{Simulated landscapes.}
I simulated several sets of landscapes using known parameters to evaluate
different statistical methods' performance. The first set of landscapes included
the three competing species shown in Figure 1. For each of 1000 replicates, I
generated a landscape with 100 sites by sampling from a probability distribution
defined by the figure's interaction coefficients (Appendix 1). Each of the
methods described below was then evaluated on its ability to correctly infer
that the two shrub species competed with one another, despite their frequent
co-occurrence in unshaded areas.

I then simulated landscapes with up to 20 interacting species at 25, 200, or
1600 sites using three increasingly complex models (50 replicates for each
combination of size and model). The simplest set of simulated landscapes were
generated with Gibbs samples from Equation 1, and provided a test of methods'
ability to isolate direct species-level interactions from a web of indirect 
ones. For each replicate, I randomly drew the “true” $\beta$ coefficient
magnitudes an exponential distribution with rate 1 so that
most species pairs interacted negligibly but a few interactions were strong
enough to propagate through the newtwork. I randomly assigned 25% of the
interactions to be positive; the remainder were negative.

@@

In the third set of landscapes, I simulated each species' abundance (instead of
just their presence/absence), and interactions between species occurred on a 
per-capita basis. To avoid runaway mutualisms leading to infinite abundance,
all interaction coefficients were negative in these simulations. If a method
performs well on these landscapes, it would indicate some robustness to the
biolocical mechanisms underlying species' interactiosn with one another.

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 3).
First, I used the rosalia package [@harris_rosalia_2015] to fit Markov network
models, as described above. For the analyses with 20 species, a weak regularizer
(equivalent to a logistic prior with scale 2) ensured that the estimates were
always finite.

I also evaluated six alternative methods: five from the existing literature,
plus a novel combination of two of these methods. The first alternative 
interaction metric was the sample correlation between species' presence-absence
vectors, which summarizes their marginal association.  Next, I used partial 
correlations, which summarize species' conditional relationships. This approach, 
which is  closely related to linear regression, is common in molecular biology
[@friedman_sparse_2008], but is rare in ecology (see @albrecht_spatial_2001 and
@faisal_inferring_2010 for two exceptions). In the context of non-Gaussian data,
the partial correlation can be thought of as a computationally efficient
approximation to the full Markov network model [@loh_structure_2013]. Partial
correlations are undefined for landscapes with perfectly-correlated species
pairs, so I used the regularized estimate provided by the corpcor package’s 
`pcor.shrink` function with the default settings [@schafer_corpcor_2014].

@@

\noindent
The results presented above show that Markov networks can reliably recover
species' pairwise interactions from species composition data, even for cases where
environmental heterogeneity and indirect interactions cause ecologists' typical
null modeling approaches to reliably fail. Partial covariances and generalized
linear models can both provide computationally efficient approximations, but
with somewhat lower accuracy [especially for typically-sized data sets with small
numbers of sites; @gotelli_empirical_2009]. The difference in accuracy may be
larger for real data sets than for the simulated landscapes in Figure 3,
however; linear approximations to the Markov network make larger errors when
the interaction matrix is structured [e.g. due to guilds or trophic levels;
@loh_structure_2013]. Similarly, the separate generalized linear models
for each species can severely overfit in some cases [@lee_learning_2012]. The
full Markov network should thus be preferred to the approximations when
it is computationally tractable.

@@

**Figure 1: A.** A small network of three competing species.
The tree (top) tends not to co-occur with either of the two shrub species, as
indicated by the strongly negative coefficient linking them. The two shrub
species also compete with one another, but more weakly (circled coefficient).
**B.** In spite of the competitive interactions between the two shrub species,
their shared tendency to occur in locations without trees makes their occurrence
vectors positively correlated (circled). **C.** Controlling for trees with a
conditional (all-else-equal) approach such as a partial covariance or a Markov
network leads to correct identification of the negative shrub-shrub interaction
(circled). See Appendix 1 and the results for "three species" for more details.

>>>>

**Abstract:** Inferring species interactions from co-occurrence data is one of
the most controversial tasks in community ecology. One difficulty is that a
single pairwise interaction can ripple through an ecological network and produce
surprising indirect consequences. For example, the negative correlation between
two competing species can be reversed in the presence of a third species that 
outcompetes both of them. Here, I apply models from statistical physics, called 
Markov networks or Markov random fields, that can predict
the direct and indirect consequences of any possible species interaction matrix.
Interactions in these models can be estimated from observed co-occurrence
rates via maximum likelihood, controlling for indirect effects. Using simulated
landscapes with known interactions, I evaluated Markov networks and six existing
approaches. The Markov networks consistently outperformed other methods, 
correctly isolating direct interactions between species pairs even when indirect
interactions or abiotic factors largely overpowered them. Two computationally 
efficient approximations, based on controlling for indirect effects with linear 
or generalized linear models, also performed well. Null models showed no 
evidence of being able to control for indirect effects, and reliably yielded 
incorrect inferences when such effects were present.

@@

\noindent \textbf{Simulated landscapes.}
I simulated several sets of landscapes using known parameters so that model
estimates could be compared with "true" values. The first set of landscapes 
included the three competing species shown in Figure 1. For each of 1000 
replicates, I generated a landscape with 100 sites by sampling from a 
probability distribution defined by the figure's interaction coefficients 
(Appendix 1). Each of the methods described below was then evaluated on its 
ability to correctly infer that the two shrub species competed with one another,
despite their frequent co-occurrence in unshaded areas.

I then simulated landscapes with up to 20 interacting species at 25, 200, or
1600 sites using three increasingly complex models (50 replicates for each
combination of size and model). The simplest set of simulated landscapes were
generated with Gibbs samples from Equation 1. For each replicate, I randomly 
drew the “true” $\beta$ coefficient magnitudes from an exponential distribution 
with rate 1 so that most species pairs interacted negligibly but a few 
interactions were strong enough to propagate through the newtwork. I randomly 
assigned 25% of the interactions to be positive; the remainder were negative.

@@

In the third set of landscapes, I simulated each species' abundance (instead of
just their presence/absence), and interactions between species occurred on a 
per-capita basis. To avoid runaway mutualisms leading to infinite abundance,
all interaction coefficients were negative in these simulations. Good 
performance on these landscapes would indicate some robustness to the
mechanistic details of species interactions.

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 3).
First, I used the rosalia package [@harris_rosalia_2015] to fit Markov network
models, as described above. For the analyses with 20 species, a 
weakly-informative regularizer (equivalent to a logistic prior with location 0
and scale 2) ensured that the estimates were always finite.

I also evaluated six alternative methods: five from the existing literature,
plus a novel combination of two of these methods. The first alternative 
interaction metric was the sample correlation between species' presence-absence
vectors, which summarizes their marginal association.  Next, I used partial 
correlations, which summarize species' conditional relationships. This approach, 
which is  closely related to linear regression, is common in molecular biology
[@friedman_sparse_2008], but is rare in ecology (see @albrecht_spatial_2001 and
@faisal_inferring_2010 for two exceptions). In the context of non-Gaussian data,
the partial correlation (or partial covariance) can be thought of as a 
computationally efficient approximation to the full Markov network model
[@loh_structure_2013]. Partial correlations are undefined for landscapes with
perfectly-correlated species pairs, so I used the regularized estimate provided 
by the corpcor package’s `pcor.shrink` function with the default settings
[@schafer_corpcor_2014].

@@

\noindent
The results presented above show that Markov networks can reliably recover
species' pairwise interactions from species composition data, even for cases where
environmental heterogeneity and indirect interactions cause ecologists' typical
null modeling approaches to reliably fail. Partial correlations and generalized
linear models can both provide computationally efficient approximations, but
with somewhat lower accuracy [especially for typically-sized data sets with small
numbers of sites; @gotelli_empirical_2009]. The difference in accuracy may be
larger for real data sets than for the simulated landscapes in Figure 3,
however; linear approximations to the Markov network make larger errors when
the interaction matrix is structured [e.g. due to guilds or trophic levels;
@loh_structure_2013]. Similarly, the separate generalized linear models
for each species can severely overfit in some cases [@lee_learning_2012]. The
full Markov network should thus be preferred to the approximations when
it is computationally tractable.

@@

**Figure 1: A.** A small network of three competing species.
The tree (top) tends not to co-occur with either of the two shrub species, as
indicated by the strongly negative coefficient linking them. The two shrub
species also compete with one another, but more weakly (circled coefficient).
**B.** In spite of the competitive interactions between the two shrub species,
their shared tendency to occur in locations without trees makes their occurrence
vectors positively correlated (circled). **C.** Controlling for trees with a
conditional (all-else-equal) approach such as a partial correlation or a Markov
network leads to correct identification of the negative shrub-shrub interaction
(circled). See Appendix 1 and the results for "three species" for more details.

### 2016-02-18 Harris_2015/manuscript-materials/draft.tex
many changes

\textbf{Abstract:} Inferring species interactions from co-occurrence
data is one of the most controversial tasks in community ecology. One
difficulty is that a single pairwise interaction can ripple through an
ecological network and produce surprising indirect consequences. For
example, the negative correlation between two competing species can be
reversed in the presence of a third species that is capable of
outcompeting both of them. Here, I apply models from statistical
physics, called Markov networks or Markov random fields, that can
predict the direct and indirect consequences of any possible species
interaction matrix. Interactions in these models can be estimated from
observed co-occurrence rates via maximum likelihood, controlling for
indirect effects. Using simulated landscapes with known pairwise
interaction strengths, I evaluated Markov networks and six existing
approaches. The Markov networks consistently outperformed other methods,
correctly isolating direct interactions between species pairs even when
indirect interactions or abiotic factors largely overpowered them. Two
computationally efficient approximations, based on controlling for
indirect effects with linear or generalized linear models, also
performed well. Indirect effects reliably caused a common null modeling
approach to produce incorrect inferences, however.

@@

\noindent where the second sum is over all \(\frac{1}{2}n(n-1)\) pairs
of \(n\) species. In this model, \(\alpha_{i}\) is an intercept term
determining the amount that the presence of species \(i\) contributes to
the log-probability of \(\vec{y}\); it directly controls the prevalence
of species \(i\). Similarly, \(\beta_{ij}\) is the amount that the
co-occurrence of species \(i\) and species \(j\) contributes to the
log-probability; it controls the conditional relationship between two
species, i.e.~the probability that they will be found together, after
controlling for the other species in the network (Figure 2A, Figure 2B).
For example, if \(\beta_{ij} = +2\), then each species' odds of
occurrence would be \(e^2\) times higher when the other one is present
(as compared with otherwise equivalent sites). The relative probability
of a presence-absence vector increases when positively-associated
species co-occur and decreases when negatively-associated species
co-occur. As a result, the model tends---all else equal---to produce
assemblages where many positively-associated species pairs co-occur and
few negatively-associated pairs do (just as an ecologist might expect).
When all else is \emph{not} equal (e.g.~Figure 1, where the presence of
one competitor is associated with release from another competitor), then
predicting species' overall co-occurrence rates would require
normalizing the relative probabilities of possible \(\vec{y}\) vectors
so that they sum to one (Figure 2B).

@@

When the number of species is larger than about 30, noralizing Equation
1 can become intractable, and researchers will either need to
approximate it (Lee and Hastie 2012) or approximate its gradient
({\textbf{???}}). Such approximations are beyond the scope of this
paper, however. For the analyses presented here, I used the rosalia
package (Harris 2015a) for the R programming language (R Core Team 2015)
to calculate \(p(\vec{y}; \alpha, \beta)\) and its gradients exactly
(Murphy 2012); the package passes these functions to the ``BFGS'' method
in R's general-purpose optimizer, which finds values for \(\alpha\) and
\(\beta\).

\noindent \textbf{Simulated landscapes.} I simulated several sets of
landscapes using known parameters to evaluate different statistical
methods' performance. The first set of landscapes included the three
competing species shown in Figure 1. For each of 1000 replicates, I
generated a landscape with 100 sites by sampling from a probability
distribution defined by the figure's interaction coefficients (Appendix
1). Each of the methods described below was then evaluated on its
ability to correctly infer that the two shrub species competed with one
another, despite their frequent co-occurrence in unshaded areas.

I then simulated landscapes with up to 20 interacting species at 25,
200, or 1600 sites using three increasingly complex models (50
replicates for each combination of size and model). The simplest set of
simulated landscapes were generated with Gibbs samples from Equation 1,
and provided a test of methods' ability to isolate direct species-level
interactions from a web of indirect ones. For each replicate, I randomly
drew the ``true'' \(\beta\) coefficient magnitudes an exponential
distribution with rate 1 so that most species pairs interacted
negligibly but a few interactions were strong enough to propagate
through the newtwork. I randomly assigned 25\% of the interactions to be
positive; the remainder were negative.

@@

In the third set of landscapes, I simulated each species' abundance
(instead of just their presence/absence), and interactions between
species occurred on a per-capita basis. To avoid runaway mutualisms
leading to infinite abundance, all interaction coefficients were
negative in these simulations. If a method performs well on these
landscapes, it would indicate some robustness to the biolocical
mechanisms underlying species' interactiosn with one another.

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 3).
First, I used the rosalia package (Harris 2015a) to fit Markov network
models, as described above. For the analyses with 20 species, a weak
regularizer (equivalent to a logistic prior with scale 2) ensured that
the estimates were always finite.

I also evaluated six alternative methods: five from the existing
literature, plus a novel combination of two of these methods. The first
alternative interaction metric was the sample correlation between
species' presence-absence vectors, which summarizes their marginal
association. Next, I used partial correlations, which summarize species'
conditional relationships. This approach, which is closely related to
linear regression, is common in molecular biology (Friedman et al.
2008), but is rare in ecology (see Albrecht and Gotelli (2001) and
Faisal et al. (2010) for two exceptions). In the context of non-Gaussian
data, the partial correlation can be thought of as a computationally
efficient approximation to the full Markov network model (Loh and
Wainwright 2013). Partial correlations are undefined for landscapes with
perfectly-correlated species pairs, so I used the regularized estimate
provided by the corpcor package's \texttt{pcor.shrink} function with the
default settings (Schäfer et al. 2014).

@@

\noindent
The results presented above show that Markov networks can reliably
recover species' pairwise interactions from species composition data,
even for cases where environmental heterogeneity and indirect
interactions cause ecologists' typical null modeling approaches to
reliably fail. Partial covariances and generalized linear models can
both provide computationally efficient approximations, but with somewhat
lower accuracy (especially for typically-sized data sets with small
numbers of sites; Gotelli and Ulrich 2009). The difference in accuracy
may be larger for real data sets than for the simulated landscapes in
Figure 3, however; linear approximations to the Markov network make
larger errors when the interaction matrix is structured (e.g.~due to
guilds or trophic levels; Loh and Wainwright 2013). Similarly, the
separate generalized linear models for each species can severely overfit
in some cases (Lee and Hastie 2012). The full Markov network should thus
be preferred to the approximations when it is computationally tractable.

@@

\textbf{Figure 1: A.} A small network of three competing species. The
tree (top) tends not to co-occur with either of the two shrub species,
as indicated by the strongly negative coefficient linking them. The two
shrub species also compete with one another, but more weakly (circled
coefficient). \textbf{B.} In spite of the competitive interactions
between the two shrub species, their shared tendency to occur in
locations without trees makes their occurrence vectors positively
correlated (circled). \textbf{C.} Controlling for trees with a
conditional (all-else-equal) approach such as a partial covariance or a
Markov network leads to correct identification of the negative
shrub-shrub interaction (circled). See Appendix 1 and the results for
``three species'' for more details.

>>>>

\textbf{Abstract:} Inferring species interactions from co-occurrence
data is one of the most controversial tasks in community ecology. One
difficulty is that a single pairwise interaction can ripple through an
ecological network and produce surprising indirect consequences. For
example, the negative correlation between two competing species can be
reversed in the presence of a third species that outcompetes both of
them. Here, I apply models from statistical physics, called Markov
networks or Markov random fields, that can predict the direct and
indirect consequences of any possible species interaction matrix.
Interactions in these models can be estimated from observed
co-occurrence rates via maximum likelihood, controlling for indirect
effects. Using simulated landscapes with known interactions, I evaluated
Markov networks and six existing approaches. The Markov networks
consistently outperformed other methods, correctly isolating direct
interactions between species pairs even when indirect interactions or
abiotic factors largely overpowered them. Two computationally efficient
approximations, based on controlling for indirect effects with linear or
generalized linear models, also performed well. Null models showed no
evidence of being able to control for indirect effects, and reliably
yielded incorrect inferences when such effects were present.

@@

\noindent where the second sum is over all \(\frac{1}{2}n(n-1)\) pairs
of \(n\) species. In this model, \(\alpha_{i}\) is an intercept term
determining the amount that the presence of species \(i\) contributes to
the log-probability of \(\vec{y}\); it directly controls the prevalence
of species \(i\). Similarly, \(\beta_{ij}\) is the amount that the
co-occurrence of species \(i\) and species \(j\) contributes to the
log-probability; it determines the conditional relationship between two
species, i.e.~the probability that they will be found together, after
controlling for the other species in the network (Figure 2A, Figure 2B).
For example, if \(\beta_{ij} = +2\), then each species' odds of
occurrence would be \(e^2\) times higher when the other one is present
(as compared with otherwise equivalent sites). The relative probability
of a presence-absence vector increases when positively-associated
species co-occur and decreases when negatively-associated species
co-occur. As a result, the model tends---all else equal---to produce
assemblages where many positively-associated species pairs co-occur and
few negatively-associated pairs do (just as an ecologist might expect).
When all else is \emph{not} equal (e.g.~Figure 1, where the presence of
one competitor is associated with release from another competitor), then
predicting species' overall co-occurrence rates would require
normalizing the relative probabilities of possible \(\vec{y}\) vectors
so that they sum to one (Figure 2B).

@@

When the number of species is larger than about 30, noralizing Equation
1 can become intractable, and researchers will either need to
approximate it (Lee and Hastie 2012) or approximate its gradient
({\textbf{???}}) if they want to fit a Markov network model. For the
analyses presented here, I used the rosalia package (Harris 2015a) for
the R programming language (R Core Team 2015) to calculate
\(p(\vec{y}; \alpha, \beta)\) and its gradients exactly (Murphy 2012);
the package passes these functions to the ``BFGS'' method in R's
general-purpose optimizer, which finds values for \(\alpha\) and
\(\beta\).

\noindent \textbf{Simulated landscapes.} I simulated several sets of
landscapes using known parameters so that model estimates could be
compared with ``true'' values. The first set of landscapes included the
three competing species shown in Figure 1. For each of 1000 replicates,
I generated a landscape with 100 sites by sampling from a probability
distribution defined by the figure's interaction coefficients (Appendix
1). Each of the methods described below was then evaluated on its
ability to correctly infer that the two shrub species competed with one
another, despite their frequent co-occurrence in unshaded areas.

I then simulated landscapes with up to 20 interacting species at 25,
200, or 1600 sites using three increasingly complex models (50
replicates for each combination of size and model). The simplest set of
simulated landscapes were generated with Gibbs samples from Equation 1.
For each replicate, I randomly drew the ``true'' \(\beta\) coefficient
magnitudes from an exponential distribution with rate 1 so that most
species pairs interacted negligibly but a few interactions were strong
enough to propagate through the newtwork. I randomly assigned 25\% of
the interactions to be positive; the remainder were negative.

@@

In the third set of landscapes, I simulated each species' abundance
(instead of just their presence/absence), and interactions between
species occurred on a per-capita basis. To avoid runaway mutualisms
leading to infinite abundance, all interaction coefficients were
negative in these simulations. Good performance on these landscapes
would indicate some robustness to the mechanistic details of species
interactions.

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 3).
First, I used the rosalia package (Harris 2015a) to fit Markov network
models, as described above. For the analyses with 20 species, a
weakly-informative regularizer (equivalent to a logistic prior with
location 0 and scale 2) ensured that the estimates were always finite.

I also evaluated six alternative methods: five from the existing
literature, plus a novel combination of two of these methods. The first
alternative interaction metric was the sample correlation between
species' presence-absence vectors, which summarizes their marginal
association. Next, I used partial correlations, which summarize species'
conditional relationships. This approach, which is closely related to
linear regression, is common in molecular biology (Friedman et al.
2008), but is rare in ecology (see Albrecht and Gotelli (2001) and
Faisal et al. (2010) for two exceptions). In the context of non-Gaussian
data, the partial correlation (or partial covariance) can be thought of
as a computationally efficient approximation to the full Markov network
model (Loh and Wainwright 2013). Partial correlations are undefined for
landscapes with perfectly-correlated species pairs, so I used the
regularized estimate provided by the corpcor package's
\texttt{pcor.shrink} function with the default settings (Schäfer et al.
2014).

@@

\noindent
The results presented above show that Markov networks can reliably
recover species' pairwise interactions from species composition data,
even for cases where environmental heterogeneity and indirect
interactions cause ecologists' typical null modeling approaches to
reliably fail. Partial correlations and generalized linear models can
both provide computationally efficient approximations, but with somewhat
lower accuracy (especially for typically-sized data sets with small
numbers of sites; Gotelli and Ulrich 2009). The difference in accuracy
may be larger for real data sets than for the simulated landscapes in
Figure 3, however; linear approximations to the Markov network make
larger errors when the interaction matrix is structured (e.g.~due to
guilds or trophic levels; Loh and Wainwright 2013). Similarly, the
separate generalized linear models for each species can severely overfit
in some cases (Lee and Hastie 2012). The full Markov network should thus
be preferred to the approximations when it is computationally tractable.

@@

\textbf{Figure 1: A.} A small network of three competing species. The
tree (top) tends not to co-occur with either of the two shrub species,
as indicated by the strongly negative coefficient linking them. The two
shrub species also compete with one another, but more weakly (circled
coefficient). \textbf{B.} In spite of the competitive interactions
between the two shrub species, their shared tendency to occur in
locations without trees makes their occurrence vectors positively
correlated (circled). \textbf{C.} Controlling for trees with a
conditional (all-else-equal) approach such as a partial correlation or a
Markov network leads to correct identification of the negative
shrub-shrub interaction (circled). See Appendix 1 and the results for
``three species'' for more details.

### 2016-02-18 Harris_2015/recovery.Rmd
as actually submitted to Ecology
many changes

[^1]: The PDF version of this document has been manually altered to omit 150
lines of output from the `corpcor` package of the form "`## Estimating optimal shrinkage intensity lambda (correlation matrix): 0.0326`"

@@

```{r, cache = FALSE, eval = TRUE}
fit_all = function(identifier){
  ######## Import ########

@@

>>>>

@@

```{r, cache = FALSE, eval = FALSE}
fit_all = function(identifier){
  ######## Import ########

@@

### Logistic prior used for the Markov network

From the R help file for the logistic distribution:

>The Logistic distribution with location = m and scale = s has... density

>> f(x) = 1/s exp((x-m)/s) (1 + exp((x-m)/s))^-2.

> It is a long-tailed distribution with mean m and variance $\pi$^2 /3 s^2.

As noted in the main text, I used location zero and scale 2 (plotted in
red below). The "true" parameter distribution for $\beta$ is drawn in black,
for reference.  These distributions have different means and the prior is
substantially wider than the distribution of true parameters (mean absolute
deviation from zero about `r round(mean(abs(rlogis(1E7, scale = 2))), 1)`
times larger).

```{r}
curve(ifelse(x < 0, .75 * dexp(abs(x)), .25 * dexp(x)), from = -15, to = 15,
      n = 1000, ylab = "Density")
curve(dlogis(x, location = 0, scale = 2), add = TRUE, col = "red")
```

The log of this density was added to the log-likelihood to calculate an
un-normalized log-posterior, which is optimized by the `optim` function in the
`stats` package.

This prior distribution is "weakly informative" in the sense of Gelman et al.
(2008. Annals of Applied Statistics, "A weakly informative default prior 
distribution for logistic and other regression models"), as shown below:
the logistic regularizer pulls the estimated values toward zero less strongly
than Gelman et al.'s weakly-informative Cauchy prior does.

```{r}
x = as.matrix(
  read.csv("fakedata/matrices/25-no_env1.csv")[,-1]
)

# Don't estimate parameters for species that were never observed
x = x[ , colMeans(x) != 0]

# Estimate parameters using a weakly informative logistic prior
logistic_fit = rosalia(x, prior = make_logistic_prior(scale = 2),
                             trace = FALSE)

# Estimate parameters using Gelman et al.'s weakly informative Cauchy prior
cauchy_fit = rosalia(x, prior = make_cauchy_prior(scale = 2.5),
                           trace = FALSE)

plot(logistic_fit$beta, cauchy_fit$beta)
abline(0,1, h = 0, v = 0, col = "darkgray")
```

### 2016-02-18 Harris_2015/recovery.tex
many changes

@@

This document describes how the different models were fit to the
simulated data from Appendix 2 and how each model's performance was
evaluated.\footnote{The PDF version of this document has been manually
  altered to omit 150 lines of output from the \texttt{corpcor} package
  of the form
  ``\texttt{\#\#\ Estimating\ optimal\ shrinkage\ intensity\ lambda\ (correlation\ matrix):\ 0.0326}''}

@@

>>>>

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

@@

This document describes how the different models were fit to the
simulated data from Appendix 2 and how each model's performance was
evaluated.{[}\^{}1{]}

@@

\subsubsection{Logistic prior used for the Markov
network}\label{logistic-prior-used-for-the-markov-network}

From the R help file for the logistic distribution:

\begin{quote}
The Logistic distribution with location = m and scale = s has\ldots{}
density
\end{quote}

\begin{quote}
\begin{quote}
f(x) = 1/s exp((x-m)/s) (1 + exp((x-m)/s))\^{}-2.
\end{quote}
\end{quote}

\begin{quote}
It is a long-tailed distribution with mean m and variance \(\pi\)\^{}2
/3 s\^{}2.
\end{quote}

As noted in the main text, I used location zero and scale 2 (plotted in
red below). The ``true'' parameter distribution for \(\beta\) is drawn
in black, for reference. These distributions have different means and
the prior is substantially wider than the distribution of true
parameters (mean absolute deviation from zero about 2.8 times larger).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{curve}\NormalTok{(}\KeywordTok{ifelse}\NormalTok{(x <}\StringTok{ }\DecValTok{0}\NormalTok{, .}\DecValTok{75} \NormalTok{*}\StringTok{ }\KeywordTok{dexp}\NormalTok{(}\KeywordTok{abs}\NormalTok{(x)), .}\DecValTok{25} \NormalTok{*}\StringTok{ }\KeywordTok{dexp}\NormalTok{(x)), }\DataTypeTok{from =} \NormalTok{-}\DecValTok{15}\NormalTok{, }\DataTypeTok{to =} \DecValTok{15}\NormalTok{,}
      \DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{"Density"}\NormalTok{)}
\KeywordTok{curve}\NormalTok{(}\KeywordTok{dlogis}\NormalTok{(x, }\DataTypeTok{location =} \DecValTok{0}\NormalTok{, }\DataTypeTok{scale =} \DecValTok{2}\NormalTok{), }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{recovery_files/figure-latex/unnamed-chunk-4-1.pdf}

The log of this density was added to the log-likelihood to calculate an
un-normalized log-posterior, which is optimized by the \texttt{optim}
function in the \texttt{stats} package.

This prior distribution is ``weakly informative'' in the sense of Gelman
et al. (2008. Annals of Applied Statistics, ``A weakly informative
default prior distribution for logistic and other regression models''),
as shown below: the logistic regularizer pulls the estimated values
toward zero less strongly than Gelman et al.'s weakly-informative Cauchy
prior does.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x =}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(}
  \KeywordTok{read.csv}\NormalTok{(}\StringTok{"fakedata/matrices/25-no_env1.csv"}\NormalTok{)[,-}\DecValTok{1}\NormalTok{]}
\NormalTok{)}

\CommentTok{# Don't estimate parameters for species that were never observed}
\NormalTok{x =}\StringTok{ }\NormalTok{x[ , }\KeywordTok{colMeans}\NormalTok{(x) !=}\StringTok{ }\DecValTok{0}\NormalTok{]}

\CommentTok{# Estimate parameters using a weakly informative logistic prior}
\NormalTok{logistic_fit =}\StringTok{ }\KeywordTok{rosalia}\NormalTok{(x, }\DataTypeTok{prior =} \KeywordTok{make_logistic_prior}\NormalTok{(}\DataTypeTok{scale =} \DecValTok{2}\NormalTok{),}
                             \DataTypeTok{trace =} \OtherTok{FALSE}\NormalTok{)}

\CommentTok{# Estimate parameters using Gelman et al.'s weakly informative Cauchy prior}
\NormalTok{cauchy_fit =}\StringTok{ }\KeywordTok{rosalia}\NormalTok{(x, }\DataTypeTok{prior =} \KeywordTok{make_cauchy_prior}\NormalTok{(}\DataTypeTok{scale =} \FloatTok{2.5}\NormalTok{),}
                           \DataTypeTok{trace =} \OtherTok{FALSE}\NormalTok{)}

\KeywordTok{plot}\NormalTok{(logistic_fit$beta, cauchy_fit$beta)}
\KeywordTok{abline}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{, }\DataTypeTok{h =} \DecValTok{0}\NormalTok{, }\DataTypeTok{v =} \DecValTok{0}\NormalTok{, }\DataTypeTok{col =} \StringTok{"darkgray"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{recovery_files/figure-latex/unnamed-chunk-5-1.pdf}
